---
title: "Coding Exercises Part 1"
author: "Paige Lewis"
date: "2025-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**PART 1**
setup
```{r}
source("setup.R")
```

load data
```{r}
data("knz_bison")
```

Create a new column in the bison dataset that calculates the individual’s age at the time of observation.

```{r}
knz_bison <- knz_bison %>%
  mutate(age = rec_year - animal_yob)
```


Write a function that will first filter the bison dataset based on any user-specified weight threshold (returning the filtered dataset).


taking median of weights, and filtering to use only median and greater - ideally filtering (to some extent) the age of the bison as well
```{r}
median(knz_bison$animal_weight, na.rm = TRUE)
```
777

```{r}
bigbois <- knz_bison %>% 
  filter(animal_weight > 777, na.rm = TRUE)
```


Create a boxplot showing weight distribution between male and female bison, and then conduct a t-test to see if weight is significantly difference between males and females.

```{r}
bigbois %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = animal_sex, y = animal_weight, fill = animal_sex))
```
